Benchmarking Python Deep Learning Frameworks\\ for Language Modeling on GPUs

### 2016-12-14 paper/main.tex
remove neural LM todo
reorganizing / fleshing out sections 4.2 and 4.3

Traditionally, neural models are trained until their performance on a \textit{development} set stops improving. Therefore, a full epoch of training (where all the training data is used) is usually followed by a predictive pass through development data.
\section{Benchmarks}
\label{sec:benchmarks}
\subsection{Time}
The first benchmark to look at is how long training takes. Table \ref{tab:timing} summarizes various breakdowns. \ms{Summarize results: ie DyNet is the best (so far) (2 sentences perhaps)}.
\begin{table}
\begin{tabular}{c|ccc}
Time					& DyNet 		& TensorFlow 	& Theano \\ \hline
Convergence		& 26h10m 		& 35h25m 			& (TBD)\\
Train epoch 		& 2h20m		& 6h57m 			& 6h35m \\
Test epoch 		& 3m4s			& 7m45s 			& 24m \\
Train batch 		& 1s 				& 5s 					& 1s \\
Test batch 			& $<$1s 		& $<$1s 			& $<$1s \\
\end{tabular}
\caption{\label{tab:timing}Total runtime on various subtasks.}
\end{table}
\subsection{Function-level metrics}
Using NVidia's built-in GPU profiler, \verb!nvprof!, we measured various ways in which the different frameworks utilized the GPU hardware, specifically focusing on the matrix multiply operation implemented in the function \verb!magma_lds128_sgemm_kernel()!. Table \ref{tab:metrics} summarizes. \ms{Summarize results: i.e. they're all in the same order of magnitude, nothing really tells us why DyNet would be much more efficient.} DyNet also spends a large amount of time in the \texttt{CUDA memset} function, unlike the other two frameworks. This may relate to DyNet's memory model: the total amount of memory available to forward and backward passes through the computation graph is specified at runtime, and DyNet may manually manage its memory use in a way TensorFlow and Theano do not.

@@

\subsection{Cuda Profiler}
\label{subsec:cudaprof}
We used the CUDA Profiler Toolkit \cite{nvprof} to characterize our implementations, specifically on a simplified task of training on one batch.

We profile which methods take the most amount of time, and look into a variety of metrics.
\begin{itemize}
\item \verb!achieved_occupancy! - number of GPU warps used vs. total number of warps available
\item \verb!sm_efficiency! - how much of the runtime is spent performing actual computation
\item \verb!warp_execution_efficiency! - how efficient is execution within a warp; number of registers per thread is likely a limiting factor for all implementations.
\item \verb!warp_nonpred_execution_efficiency! - how efficient is execution within a warp on non-predicated (i.e. non-branching) instructions.
\item \verb!dram_read_throughput! - read throughput from RAM to GPU
\item \verb!dram_write_throughput! - write throughput from GPU to RAM
\end{itemize}

>>>>

Traditionally, neural models are trained until their performance on a \textit{development} set stops improving. Therefore, a full epoch of training (where all the training data is used) is usually followed by a predictive pass through development data.
\section{Benchmarks}
\label{sec:benchmarks}
\subsection{Time}
The first benchmark to look at is how long training takes. Table \ref{tab:timing} summarizes various breakdowns. \ms{Summarize results: ie DyNet is the best (so far) (2 sentences perhaps)}.
\begin{table}
\begin{tabular}{c|ccc}
Time					& DyNet 		& TensorFlow 	& Theano \\ \hline
Convergence		& 26h10m 		& 35h25m 			& (TBD)\\
Train epoch 		& 2h20m		& 6h57m 			& 6h35m \\
Test epoch 		& 3m4s			& 7m45s 			& 24m \\
Train batch 		& 1s 				& 5s 					& 1s \\
Test batch 			& $<$1s 		& $<$1s 			& $<$1s \\
\end{tabular}
\caption{\label{tab:timing}Total runtime on various subtasks.}
\end{table}

\subsection{Kernel function usage}
Using NVidia's built-in GPU profiler, \verb!nvprof!, we first looked into which kernel function calls were most heavily utilized by profiling training on a single batch. As can be seen in Table~\ref{tab:pcttime}, a large portion of the runtime is spent on matrix multiplication (specifically, a variant of \texttt{\detokenize{magma_lds_128_sgemm_kernel()}}) for all frameworks; this is expected given how neural networks work.

Unlike the other two frameworks, DyNet also spends a large amount of time in the \texttt{CUDA memset} function. This may relate to DyNet's memory model: the total amount of memory available to forward and backward passes through the computation graph is specified at runtime, and DyNet may manually manage its memory use in a way TensorFlow and Theano do not.

@@

\subsection{CUDA profiler metrics}
\label{subsec:cudaprof}

Given the high usage of the matrix multiply operation \verb!magma_lds128_sgemm_kernel()!, we
then looked into a variety of CUDA profiler metrics for this operation.

We collected data for the following metrics:
\begin{itemize}
\item \verb!achieved_occupancy! -- number of GPU warps used vs.\ total number of warps available.
\item \verb!sm_efficiency! -- how much of the runtime is spent performing actual computation.
\item \verb!warp_execution_efficiency! -- how efficient execution is within a warp; note that the number of registers per thread is likely a limiting factor for all implementations.
\item \verb!warp_nonpred_execution_efficiency! -- how efficient execution is within a warp on non-predicated (i.e. non-branching) instructions.
\item \verb!dram_read_throughput! -- read throughput from RAM to GPU.
\item \verb!dram_write_throughput! -- write throughput from GPU to RAM.
\end{itemize}

Table \ref{tab:metrics} summarizes the results on a single batch. We found that all of the measurements were within similar orders of magnitude; nothing at this level directly points to why DyNet performance was considerably faster than the other frameworks.
